Towards a free/open-source universal-dependency treebank for Kazakh

### 2015-06-20 2015-turklang-kazdep/paper.tex
section
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60911 72bbbca6-d526-0410-a7d9-f06f51895060

\section{Annotation guidelines}\label{sec:annotation}

% general procedure

% - disambiguate
% - annotate
% - check the tree structure
% - fix errors
% - check for label consistency
% - fix errors

>>>>

\section{Annotation}\label{sec:annotation}

Annotation of the treebank followed an iterative approach. The first stage was to
select a text and preprocess it using the morphological analyser and constraint grammar.
The output of this process was then manually disambiguated by choosing the most appropriate
analysis in context, or adding a new analysis if one was not returned by the morphological
analyser. This latter step was mostly in the case of unknown proper nouns.\footnote{New lexemes
found during the development of the corpus will contribute to the development of the
morphological analyser.}

After the text was disambiguated, it was annotated by a single annotator for dependency
structure relying on the annotation guidelines for universal dependencies.\footnote{\url{http://universaldependencies.github.io/docs/u/overview/syntax.html}}

Where it was not clear how to annotate a particular structure, it was left pending, that
is marked with the {\sc dep} relation. After a text was annotated, graphical representations
of the trees were produced and proofread by a second annotator. The final analysis for a
sentence was decided on during discussions between the two annotators and in some cases
through the universal-dependency issue tracker.\footnote{\url{https://github.com/UniversalDependencies/docs/issues/}}

During the course of annotation, language-specific annotation guidelines for Kazakh are
being produced and published online.\footnote{\url{http://wiki.apertium.org/wiki/Dependency_parsing_for_Turkic}} In 
the remainder of this section we describe a number of Kazakh-specific decisions.
